---
title: "scRNA-Seq Anlaysis - 05 - Clustering"
output:
  github_document:
    toc: true
---

# Dependencies
```{r}
library(scater)
library(SC3)
library(dplyr)
library(tidyr)
library(ggforce)
```

# Load the data
```{r}
sce <- readRDS("../output/sce.normalized.imputed.rds")
```

# Clustering with SC3

## Estimating optimal k
```{r}
sce.temp <- sce
exprs(sce.temp) <- assay(sce.temp, "exprs.magic")
sce.temp <- sc3_estimate_k(sce.temp)
```

```{r}
metadata(sce.temp)
```
Looks like optimal k for clustering is 3.

## Clustering
```{r}
rowData(sce.temp)$feature_symbol <- rownames(sce.temp) # I should actually make this symbols
sce.temp <- sc3(sce.temp, ks=2:3, biology=TRUE, rand_seed=100)
```

Bringing the results back to the main sce
```{r}
colData(sce) <- colData(sce.temp)
rowData(sce) <- rowData(sce.temp)
```

# Assessing Clustersing
## PCA

```{r, fig.width=4.75, fig.height=3.5}
plotPCA(sce, exprs_values="exprs.magic", shape_by="Condition",
        colour_by="sc3_2_clusters")
plotPCA(sce, exprs_values="exprs.magic", shape_by="Condition",
        colour_by="sc3_3_clusters")
```

### Figure
```{r, fig.width=4.75, fig.height=3.5}
cols <- c("#44B3C2", "#F1A94E", "#E45641", "#5D4C46", "#7B8D8E", "#F2EDD8")
sce <- plotPCA(sce, exprs_values="exprs.magic", return_SCE=T)
pca <- as.data.frame(sce@reducedDims$PCA[,1:2])
pca$Condition <- factor(sce$Condition)
pca$Cluster <- factor(sce$sc3_3_clusters)

#3 clusters
pca.3clust <- ggplot(pca, aes(x=PC1, y=PC2)) +
  geom_point(size=0.3, aes(colour=Cluster, shape=Condition)) +
  xlab("PC1") + ylab("PC2") +
  scale_color_manual(values=c("grey40", cols[3], cols[1])) +
  theme_classic() +
  theme(axis.text = element_text(size=5, color="black"),
        axis.title.x = element_text(size=6, color="black", 
                                    margin=unit(c(1,0,0,0), "mm")),
        axis.title.y = element_text(size=6, color="black", 
                                    margin=unit(c(0,1,0,0), "mm")),
        axis.line = element_line(size=0.4, color="black"),
        axis.ticks = element_line(size=0.4, color="black"),
        legend.text = element_text(size=6),
        legend.title = element_text(size=6),
        legend.key.size = unit(0.45, 'lines'),
        legend.margin=margin(c(-4,0,0,-5), "mm"),
        plot.background=element_rect(fill="transparent", colour=NA))
pca.3clust
ggsave(pca.3clust, file="../figs/pca.3clust.pdf", width=1.9, height=1.215,
       useDingbats=F)

#2 clusters
pca$Cluster <- as.numeric(sce$sc3_2_clusters)
pca$Cluster[pca$Cluster==1] <- 3 #Stupid round-about way to swap cluster IDs
pca$Cluster[pca$Cluster==2] <- 1
pca$Cluster[pca$Cluster==3] <- 2
pca$Cluster <- factor(pca$Cluster)
pca.2clust <- ggplot(pca, aes(x=PC1, y=PC2)) +
  geom_point(size=0.3, aes(colour=Cluster, shape=Condition)) +
  xlab("PC1") + ylab("PC2") +
  scale_color_manual(values=c("grey40", cols[2])) +
  theme_classic() +
  theme(axis.text = element_text(size=5, color="black"),
        axis.title.x = element_text(size=6, color="black", 
                                    margin=unit(c(1,0,0,0), "mm")),
        axis.title.y = element_text(size=6, color="black", 
                                    margin=unit(c(0,1,0,0), "mm")),
        axis.line = element_line(size=0.4, color="black"),
        axis.ticks = element_line(size=0.4, color="black"),
        legend.text = element_text(size=6),
        legend.title = element_text(size=6),
        legend.key.size = unit(0.45, 'lines'),
        legend.margin=margin(c(-4,0,0,-5), "mm"),
        plot.background=element_rect(fill="transparent", colour=NA))
pca.2clust
ggsave(pca.2clust, file="../figs/pca.2clust.pdf", width=1.9, height=1.215,
       useDingbats=F)
```

# Proportions of each condition in each cluster

Two clusters splits the left-right clustering pattern apart, and three clusters seems to split the left cluster near-perfectly by condition. Let's make some plots to just see the proportions.

```{r fig.width=1.7, fig.height=0.65}
clust2.prop <- as.data.frame(colData(sce)) %>%
  group_by(Condition, sc3_2_clusters) %>%
  dplyr::summarize(count = n()) %>%
  mutate(freq = (count / sum(count)) * 100)
clust2.prop$sc3_2_clusters <- as.numeric(clust2.prop$sc3_2_clusters)
clust2.prop$sc3_2_clusters[clust2.prop$sc3_2_clusters==1] <- 3 #Same stupid hack
clust2.prop$sc3_2_clusters[clust2.prop$sc3_2_clusters==2] <- 1
clust2.prop$sc3_2_clusters[clust2.prop$sc3_2_clusters==3] <- 2
clust2.prop$sc3_2_clusters <- factor(clust2.prop$sc3_2_clusters)
clust2 <- ggplot(clust2.prop, aes(x=Condition, y=freq, 
                                  fill=factor(sc3_2_clusters))) +
  geom_bar(color="black", stat="identity", size=0.3,
           position=position_stack(reverse=T)) +
  xlab("") + ylab("% of cells") +
  scale_y_continuous(limits=c(0,100), expand=c(0,0)) +
  scale_fill_manual(values=c("grey40", cols[2]), name="Cluster") +
  theme_classic() +
  theme(axis.text = element_text(size=5, color="black"),
        axis.title.x = element_text(size=6, color="black", 
                                    margin=unit(c(1,0,0,0), "mm")),
        axis.title.y = element_text(size=6, color="black", 
                                    margin=unit(c(0,1,0,0), "mm")),
        axis.line.x = element_line(size=0.4, color="black"),
        axis.line.y = element_blank(),
        axis.ticks.x = element_line(size=0.4, color="black"),
        axis.ticks.y = element_blank(),
        legend.text = element_text(size=6),
        legend.title = element_text(size=6),
        legend.key.size = unit(0.45, 'lines'),
        legend.margin=margin(c(0,0,0,-2), "mm"),
        plot.background=element_rect(fill="transparent", colour=NA)) +
  coord_flip()
clust2
ggsave(clust2, file="../figs/clusters.2.proportions.pdf", width=1.7, height=0.65,
       useDingbats=F)
```

###Three clusters
```{r, fig.width=1.7, fig.height=0.65}
clust3.prop <- as.data.frame(colData(sce)) %>%
  group_by(Condition, sc3_3_clusters) %>%
  dplyr::summarize(count = n()) %>%
  mutate(freq = (count / sum(count)) * 100)
clust3.prop <- bind_rows(clust3.prop, clust3.add)
clust3 <- ggplot(clust3.prop, aes(x=Condition, y=freq, 
                                  fill=factor(sc3_3_clusters))) +
  geom_bar(color="black", stat="identity", size=0.3,
           position=position_stack(reverse=T)) +
  xlab("") + ylab("% of cells") +
  scale_y_continuous(limits=c(0,100), expand=c(0,0)) +
  scale_fill_manual(values=c("grey40", cols[3], cols[1]), name="Cluster") +
  theme_classic() +
  theme(axis.text = element_text(size=5, color="black"),
        axis.title.x = element_text(size=6, color="black", 
                                    margin=unit(c(1,0,0,0), "mm")),
        axis.title.y = element_text(size=6, color="black", 
                                    margin=unit(c(0,1,0,0), "mm")),
        axis.line.x = element_line(size=0.4, color="black"),
        axis.line.y = element_blank(),
        axis.ticks.x = element_line(size=0.4, color="black"),
        axis.ticks.y = element_blank(),
        legend.text = element_text(size=6),
        legend.title = element_text(size=6),
        legend.key.size = unit(0.45, 'lines'),
        legend.margin=margin(c(0,0,0,-2), "mm"),
        plot.background=element_rect(fill="transparent", colour=NA)) +
  coord_flip()
clust3
ggsave(clust3, file="../figs/clusters.3.proportions.pdf", width=1.7, height=0.65,
       useDingbats=F)
```

#Cluster markers
SC3 uses an AUROC approach to define unique markers of each cluster

```{r fig.width=1.25, fig.height=1.1}
auroc <- data.frame(auroc=rowData(sce)$sc3_3_markers_auroc,
                    gene=rownames(sce),
                    cluster=rowData(sce)$sc3_3_markers_clusts)
auroc <- auroc[order(auroc$auroc),]
auroc <- na.omit(auroc)
auroc$rank <- nrow(auroc):1
rownames(auroc) <- auroc$gene
auroc$size <- 1
auroc$color <- "a"
auroc["ENSMUSG00000002020",]$size <- 2 #Ltbp2
auroc["ENSMUSG00000002020",]$color <- "b" #Ltbp2
auroc["ENSMUSG00000017969",]$size <- 2 #Ptgis
auroc["ENSMUSG00000017969",]$color <- "c" #Ptgis
auroc["ENSMUSG00000026185",]$size <- 2 #Igfbp5
auroc["ENSMUSG00000026185",]$color <- "d" #Igfbp5

auroc.plot <- ggplot(auroc, aes(x=rank, y=auroc)) +
  geom_point(data=auroc[which(auroc$size != 2),], size=0.01, aes(color=color)) +
  geom_point(data=auroc[which(auroc$size == 2),], size=1, aes(color=color)) +
  scale_color_manual(values=c("grey20", cols[2], cols[3], cols[1])) +
  xlab("Gene rank") + ylab("AUROC") +
  theme_classic() +
  theme(legend.position="none",
        axis.text.y = element_text(size=5, color="black"),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size=6, color="black", 
                                    margin=unit(c(1,0,0,0), "mm")),
        axis.title.y = element_text(size=6, color="black", 
                                    margin=unit(c(0,1,0,0), "mm")),
        axis.line = element_line(size=0.4, color="black"),
        axis.ticks.y = element_line(size=0.4, color="black"),
        axis.ticks.x = element_blank(),
        plot.background=element_rect(fill="transparent", colour=NA))
auroc.plot
ggsave(auroc.plot, file="../figs/auroc.plot.pdf", width=1.25, height=1.1,
       useDingbats=F)
```

## Marker expression across clusters
```{r}
marker.exp <- data.frame(ltbp2 = assay(sce,
                                       "exprs.magic")["ENSMUSG00000002020",],
                         ptgis = assay(sce,
                                       "exprs.magic")["ENSMUSG00000017969",],
                         igfbp5 = assay(sce,
                                       "exprs.magic")["ENSMUSG00000026185",],
                         cluster = sce$sc3_3_clusters)

marker.exp <- gather(marker.exp, "gene", "expression", -cluster)
marker.exp$gene <- factor(marker.exp$gene, levels=c("ltbp2", "ptgis", "igfbp5"))
labels <- c(ltbp2="Ltbp2", ptgis="Ptgis", igfbp5="Igfbp5")

marker.plot <- ggplot(marker.exp, aes(x=factor(cluster), y=expression)) +
  geom_sina(size=0.1, alpha=0.75, aes(colour=factor(cluster))) +
  facet_wrap(~gene, scales="free_y", labeller=labeller(gene=labels)) +
  scale_colour_manual(values=c("grey40", cols[3], cols[1])) +
  scale_y_continuous(expand=c(0,0), limits=c(0,NA)) +
  xlab("Cluster") + ylab("Log(Counts + 1)") +
  theme_classic() +
  theme(axis.title=element_text(size=5),
        axis.text=element_text(size=5, color="black"),
        axis.ticks=element_line(color="black", size=0.1),
        axis.line=element_line(color="black", size=0.1),
        axis.title.x = element_text(size=6, color="black", 
                                    margin=unit(c(1,0,0,0), "mm")),
        axis.title.y = element_text(size=6, color="black", 
                                    margin=unit(c(0,1,0,0), "mm")),
        strip.background=element_blank(),
        strip.text=element_text(size=7, color="black", margin=unit(c(0,0,0,0), "mm")),
        legend.position="none")
marker.plot
ggsave(marker.plot, file="../figs/marker.plot.pdf", width=3.4, height=1.3)
```

#Save Point
```{r}
saveRDS(sce, file="../output/sce.clustered.rds")
```

